A Typed Programmatic Interface to Contracts on the Blockchain

### 2021-02-16 papers/execution-model/model.tex
more motivation

As a bidder, you may want to implement a bidding
strategy that starts at a certain amount and increases the bid until a
limit is reached. Of course, you only want to increase if
someone else placed a higher bid. So you also want to write a
program to implement this strategy.

Neither task can be implemented as a smart contract alone because they react
to external triggers. Closing requires to schedule the activation of a
contract at a certain time and bidding 
requires to watch the current highest bid of the contract and
react if another bidder places a higher bid.

For concreteness, Listing~\ref{lst:auction-contract-header} shows the
header of an auction contract in Michelson \cite{michelson}. The
\lstinline/parameter/ clause specifies the contract's parameter
type. It is a sum type (indicated by \lstinline/or/) and each
alternative constitutes an entrypoint, named \lstinline/%close/ and
\lstinline/%bid/. Entrypoint \lstinline/%close/ takes a \lstinline{unit}
parameter and \lstinline/%bid/ takes an amount of tokens of type \lstinline{mutez}.
The \lstinline/storage/ clause declares the state of the contract,
which is a nested pair type indicating whether bidding is allowed (\lstinline{bool}), the address of the
contract owner (to prohibit unauthorized calls to \lstinline/%close/),
the highest bid (of type \lstinline{mutez}), and the bidder's address.
\begin{lstlisting}[language=michelson,numbers=none,float,caption={Header of the auction contract},label={lst:auction-contract-header},captionpos=b,emph={close,bid},emphstyle=\underbar]
parameter (or (unit %close)
              (mutez %bid)); # participant's bid
storage (pair
          (pair bool address)  # bidding allowed, contract owner
          (pair mutez address) # highest bid, bidder's address
        );
\end{lstlisting}

Here is an outline of the implementation of the entrypoints. The
\lstinline/%close/ entrypoint first checks its sender's address
against the owner's address in the store. Then it transfers the funds
to the owner, closes the contract, and trusts the owner to deliver the
goods.\footnote{For simplicity we elide safeguarding by a third-party oracle.}
The \lstinline/%bid/ entrypoint rejects and reimburses each bid that is not higher
than the existing highest bid. Otherwise, it keeps the funds
transferred, returns the previous highest bid to its owner, and stores
the current bid as the new highest bid.

@@

First the code that creates an auction contract and closes it after 48 hours. 
\begin{lstlisting}[language=Caml,numbers=none]
let initial_store = ((true, my_address), (1, my_address))
let contract = Cl.originate(auction_script, 1, initial_store)
Time.sleep(48 * 3600); (* 48h in seconds *)
Cl.call_contract(contract, left ());
\end{lstlisting}
We assume that \lstinline/Cl/ is the contract library,  \lstinline/originate/ deploys its argument contract along with
the provided funds and the \lstinline/initial_store/, 
\lstinline{auction_script} is the contract from Listing~\ref{lst:auction-contract-header}, and 
\lstinline{call_contract} invokes a contract with a parameter.

To implement the bidding strategy, we first check the state of the
contract to find the current highest bid. As long as bidding is
allowed and the current bid is below our \lstinline/limit/, we update
our bid by a given amount \lstinline/step/, and then keep watching the state of the contract by
polling it every five minutes. 
\begin{lstlisting}[numbers=none]
let rec poll (limit, step, ((bidding, _), (high_bid, _))) =
  if bidding and high_bid < limit then
    Cl.call_contract (contract, 
      right (min (high_bid + step, limit)));
    Time.sleep(5);
    let state = Cl.get_state (contract);
    poll (limit, step, state)
in
poll (limit, step, Cl.get_state (contract))
\end{lstlisting}
The \lstinline{get_state} function queries data on
the blockchain and returns the storage of a contract.

On another level, running a contract requires to propose a fee to the
blockchain, which may or may not be accepted. In any case, starting a
contract may timeout for a range of different reasons. So just
starting a contract with a fixed fee does not guarantee the contract's execution.
This scenario can also be addressed by a program that tries to run the
contract with successively increasing fees until the blockchain
accepts the contract invocation. The program can also calculate the
required fee by observing the current state of the blockchain.

(From this scenario we see the required functionality for a
programmatic interface to a blockchain:)

>>>>

As a bidder, you may want to implement a bidding
strategy that starts at a certain amount and increases the bid until a
limit is reached. Of course, you only want to increase your bid if
someone else placed a higher bid. So you also want to write a
program to implement this strategy.

Neither task can be implemented as a smart contract alone because both react
to external triggers. Closing requires to schedule the invocation of a
contract at a certain time. Bidding 
requires to watch the current highest bid of the contract and
react if another bidder places a higher bid.

For concreteness, Listing~\ref{lst:auction-contract-header} shows the
header of an auction contract in Michelson \cite{michelson}. The
\lstinline/parameter/ clause specifies the contract's parameter
type. It is a sum type (indicated by \lstinline/or/) and each
alternative constitutes an entrypoint, named \lstinline/%close/ and
\lstinline/%bid/. The caller selects the entrypoint by injecting the argument into the
\lstinline/left/ or \lstinline/right/ summand. Entrypoint \lstinline/%close/ takes a \lstinline{unit}
parameter and \lstinline/%bid/ takes an amount of tokens of type \lstinline{mutez}.
The \lstinline/storage/ clause declares the state of the contract,
which is a nested pair type indicating whether bidding is allowed (\lstinline{bool}), the address of the
contract owner (to prohibit unauthorized calls to \lstinline/%close/),
the highest bid (of type \lstinline{mutez}), and the bidder's address.
\begin{lstlisting}[language=michelson,numbers=none,float,caption={Header of the auction contract},label={lst:auction-contract-header},captionpos=b,emph={close,bid},emphstyle=\underbar]
parameter (or (unit %close)
              (mutez %bid)); # participant's bid
storage (pair
          (pair bool address)  # bidding allowed, contract owner
          (pair mutez address) # highest bid, bidder's address
        );
\end{lstlisting}

To avoid too much detail, we just outline the implementation of the entrypoints. The
\lstinline/%close/ entrypoint first checks its sender's address
against the owner's address in the store. Then it transfers the funds
to the owner, closes the contract by clearing the bidding flag, and
leaves it to the owner to deliver the
goods.\footnote{For simplicity we elide safeguarding by a third-party oracle.}
The \lstinline/%bid/ entrypoint immediately returns each bid that is not higher
than the existing highest bid. Otherwise, it keeps the funds
transferred, returns the previous highest bid to its owner, and stores
the current bid as the new highest bid.

@@

First the code that creates an auction contract and closes it after 48 hours. 
\begin{lstlisting}[language=Caml,numbers=none,basicstyle=\small]
let initial_store = ((true, my_address), (1, my_address))
let contract = Cl.originate(auction_script, 1, initial_store)
Time.sleep(48 * 3600);               (* 48h in seconds *)
Cl.call_contract(contract, left ()); (* entrypoint %close *)
\end{lstlisting}
We assume that \lstinline/Cl/ is the module containing the contract
library,  \lstinline/originate/ deploys its argument contract along
with the provided funds and the \lstinline/initial_store/, 
\lstinline{auction_script} is the contract from Listing~\ref{lst:auction-contract-header}, and 
\lstinline{call_contract} invokes a contract with a parameter.

To implement the bidding strategy, we first check the state of the
contract to find the current highest bid. As long as bidding is
allowed and the current bid is below our \lstinline/limit/, we update
our bid by a given amount \lstinline/step/, and then keep watching the state of the contract by
polling it every five minutes. 
\begin{lstlisting}[numbers=none,basicstyle=\small]
let rec poll (limit, step, ((bidding, _), (high_bid, _))) =
  if bidding and high_bid < limit then
    Cl.call_contract (contract, 
      right (min (high_bid + step, limit))); (* entrypoint %bid *)
    Time.sleep(5 * 60);
    let state = Cl.get_state (contract);
    poll (limit, step, state)
in
poll (limit, step, Cl.get_state (contract))
\end{lstlisting}
The \lstinline{get_state} function queries data on
the blockchain and returns the storage of a contract.

This code is idealized in several respects. Originating or running a contract
requires to propose a fee to the blockchain, which may or may not be
accepted.
Starting a contract may also time out for a variety of reasons. So just
invoking a contract with a fixed fee does not guarantee the contract's
execution.  Even if the invocation is locally accepted, it still takes
a couple of cycles before we can be sure the invocation is globally
accepted and incorporated in the blockchain.
Hence, after starting the invocation, we have to observe the fate of
this invocation. If it does not get incorporated, then we need to
analyse the reason and react accordingly. For example, if the
invocation was rejected because of a insufficient fee, we might want
to restart with an increased fee. Or we might decide to wait until the
invocation goes through with the low fee, if it is not that important.

Hence, we would implement a scheme similar to
the bidding strategy: start with a low fee and increase until the
contract is accepted or a fee cap is reached. On the other hand, an
observer function like \lstinline/get_state/ always succeeds.

The low-level interface that we propose in this paper requires 
(From this scenario we see the required functionality for a
programmatic interface to a blockchain:)
